% $LAAS: robotpkg.tex 2010/11/17 18:22:18 mallet $
%
% Copyright (c) 2009-2010 LAAS/CNRS
% All rights reserved.
%
% Permission to use, copy, modify, and distribute this software for any purpose
% with or without   fee is hereby granted, provided   that the above  copyright
% notice and this permission notice appear in all copies.
%
% THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES WITH
% REGARD TO THIS  SOFTWARE INCLUDING ALL  IMPLIED WARRANTIES OF MERCHANTABILITY
% AND FITNESS. IN NO EVENT SHALL THE AUTHOR  BE LIABLE FOR ANY SPECIAL, DIRECT,
% INDIRECT, OR CONSEQUENTIAL DAMAGES OR  ANY DAMAGES WHATSOEVER RESULTING  FROM
% LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT, NEGLIGENCE OR
% OTHER TORTIOUS ACTION,   ARISING OUT OF OR IN    CONNECTION WITH THE USE   OR
% PERFORMANCE OF THIS SOFTWARE.
%
%                                             Anthony Mallet on Sat Jan 10 2009
%

\documentclass[a4paper,11pt]{book}
\newif\iftth\iftth
   \input{share/robotpkg.tth}
\else
   \usepackage[T1]{fontenc}
   \usepackage{robotpkg}
\fi

\title{A guide to robotpkg}
\author{Anthony Mallet --- {\tt anthony.mallet@laas.fr}}
\date{\today}

\def\robotpkg{{\tt robotpkg} }

\begin{document} % ---------------------------------------------------------

\frontmatter
\maketitle

{\small\parindent0pt
Copyright \copyright 2006-2010 LAAS/CNRS.\\
Copyright \copyright 1997-2010 The NetBSD Foundation, Inc.\\
All rights reserved.\\

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions
are met:

\begin{enumerate}
\item Redistributions of source code must retain the above copyright
      notice, this list of conditions and the following disclaimer.

\item Redistributions in binary form must reproduce the above copyright
      notice, this list of conditions and the following disclaimer in the
      documentation and/or other materials provided with the distribution.
\end{enumerate}

THIS SOFTWARE IS PROVIDED BY THE NETBSD FOUNDATION, INC. AND CONTRIBUTORS
``AS IS'' AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED
TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
PURPOSE ARE DISCLAIMED.  IN NO EVENT SHALL THE FOUNDATION OR CONTRIBUTORS
BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
POSSIBILITY OF SUCH DAMAGE.
}


\tableofcontents
\mainmatter

\chapter{Introduction}
\label{chapter:introduction}
\input{introduction}

\chapter{The robotpkg user's guide}
\label{chapter:user}

Basically, there are two ways of using robotpkg.  The  first is to only install
the  package tools and to  use binary packages that  someone else has prepared.
The second way is  to install the  programs from source. Then  you are  able to
build your own packages,  and you can  still use  binary packages from  someone
else. Sections in this document will detail both approaches where appropriate.

\input{getting}
\input{bootstrapping}
\input{using}
\input{configuring}

\chapter{The robotpkg developer's guide}
\label{chapter:developer}

This part of the documentation deals with creating and modifying packages.

\input{pkgvars}
\input{fixing}

\chapter{The robotpkg infrastructure internals}
\label{chapter:internal}

\end{document} % -----------------------------------------------------------
